---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data
Step 1: Read in the data
```{r}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE)
unzip("activity.zip", junkpaths = TRUE)
activity <- read.csv("activity.csv")
```


## What is mean total number of steps taken per day?

```{r}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(ggplot2)
a<-activity%>%
        filter(is.na(steps) == FALSE)%>%
        group_by(date)%>%
        summarize(steps.per.day = sum(steps))
qplot(a$steps.per.day, geom = "histogram", binwidth = 1000, main = "steps per day")
```


## What is the average daily activity pattern?

3. Mean and median number of steps taken each day
```{r}
print(c("Mean steps per day: ", round(mean(a$steps.per.day), 0), 
      " Median steps per day: ", median(a$steps.per.day)))
```
4. Time series plot of the average number of steps taken
```{r}
b <- activity%>%
        filter(is.na(steps) == FALSE)%>%
        group_by(interval)%>%
        summarize(avg.steps = mean(steps))
ggplot(b, aes(x=interval, y=avg.steps))+
        geom_line()
```
See the variance in the intervals by this plot:
```{r}
ggplot(activity, aes(interval, steps)) +
  geom_point() +
  geom_line(data = b, aes(interval, avg.steps), size = 2, color = "red")
```
5. The 5-minute interval that, on average, contains the maximum number of steps
```{r}
b<- b %>%
        arrange(desc(avg.steps))
b[1,]
```

## Imputing missing values
6. Code to describe and show a strategy for imputing missing data  

```{r}
c <- filter(activity, is.na(steps)) # detect all the missing values
c <- c%>%
        left_join(b, by= "interval") # merge c(missing values) with b(averages) by Interval
activity_complete <- activity%>% # merge activity master table with averages
        left_join(select(c, c(interval, date, avg.steps)), by = c("interval", "date"))
# chose averages where missing
activity_complete$steps[is.na(activity_complete$steps)] <-activity_complete$avg.steps[is.na(activity_complete$steps)]
activity_complete <- activity_complete[,1:3] # delete redundant column

```
7. Histogram of the total number of steps taken each day after missing values are imputed
```{r}
a<-activity_complete%>%
        group_by(date)%>%
        summarize(steps.per.day = sum(steps))
qplot(a$steps.per.day)
```



## Are there differences in activity patterns between weekdays and weekends?
8. Panel plot comparing the average number of steps taken per 5-minute interval across weekdays and weekends

```{r}
d <- activity_complete%>% # insert the criteria weekend vs. week, and summarize the average
        mutate(date = as.Date(date))%>%
        mutate(wke = ifelse(weekdays(date)=="Samstag" | 
                              weekdays(date) == "Sonntag", 1,0))%>%
        group_by(interval, wke)%>%
        summarize(avg.steps = mean(steps))
        
p1 <- qplot(interval, avg.steps, data = filter(d, wke==0), geom = "line", main = "steps during week")
p2 <- qplot(interval, avg.steps, data = filter(d, wke==1), geom = "line", main = "steps during weekend")
library(ggpubr)
ggarrange(p1,p2, ncol = 1, nrow = 2)

```